---
title: "R Notebook"
output: html_notebook
---


## Prepare data from HTML file

Here is an example of preparing details from a html file downloaded from Domain.com. The html file represnts (the firs 20 results) of a search on all Tasmanian properties >10Ha.

The goal is to prepare the data in tabular format and also determine the spatial location based on the address from the listing details.

A better way will be to use Domain API to return the listing data (I am investigating this) - Kane

```{r}
# --- Setup --------------------------------------------------------------
# install.packages(c("rvest","dplyr","stringr","tidyr","readr","purrr","tidygeocoder"))
library(rvest)
library(dplyr)
library(stringr)
library(tidyr)
library(readr)
library(purrr)
library(tidygeocoder)

# --- Path to your saved HTML file --------------------------------------
file_path <- "./dataset/289 Properties for Sale in tas _ Domain.html"  # <- change if needed

# --- Helpers ------------------------------------------------------------

# Parse the mixed address string into columns
# Handles forms like: "401 Bresnehans Road, LITTLE SWANPORT TAS 7190"
# and "401 Bresnehans Road,LITTLE SWANPORTTAS7190"
parse_address <- function(addr) {
  addr <- str_squish(addr %||% "")
  if (addr == "") return(tibble(street = NA, suburb = NA, state = NA, postcode = NA))

  # First try "street, suburb state postcode"
  m1 <- str_match(addr, "^(.*?),\\s*(.*?)\\s+(TAS)\\s*(\\d{4})$")
  if (!any(is.na(m1))) {
    return(tibble(
      street   = str_squish(m1[,2]),
      suburb   = str_to_title(str_squish(m1[,3])),
      state    = str_squish(m1[,4]),
      postcode = str_squish(m1[,5])
    ))
  }

  # Then try "street,suburbSTATEPOSTCODE" (no spaces)
  m2 <- str_match(addr, "^(.*?),(.*?)(TAS)(\\d{4})$")
  if (!any(is.na(m2))) {
    return(tibble(
      street   = str_squish(m2[,2]),
      suburb   = str_to_title(str_squish(m2[,3])),
      state    = str_squish(m2[,4]),
      postcode = str_squish(m2[,5])
    ))
  }

  # Fall back: return raw in street
  tibble(street = addr, suburb = NA, state = NA, postcode = NA)
}

# Convert price text to numeric dollars where possible
# Examples:
# "$1,295,000" -> 1295000
# "Offers Over $875k" -> 875000
# "EOI" / "Expressions of Interest" -> NA
parse_price_numeric <- function(price_text) {
  if (is.na(price_text) || str_trim(price_text) == "") return(NA_real_)
  txt <- str_to_upper(gsub(",", "", price_text))

  # Skip expressions of interest, contact agent, auction guides without numbers
  if (str_detect(txt, "EOI|EXPRESSION|CONTACT|TENDER|AUCTION|BY NEGOTIATION")) return(NA_real_)

  m <- str_match(txt, "\\$?\\s*([0-9]+(?:\\.[0-9]+)?)\\s*([KM])?")
  if (any(is.na(m))) return(NA_real_)

  val <- as.numeric(m[,2])
  suf <- m[,3]

  if (!is.na(suf)) {
    if (suf == "K") val <- val * 1e3
    if (suf == "M") val <- val * 1e6
  }
  as.numeric(round(val))
}

`%||%` <- function(a, b) if (is.null(a)) b else a

# --- Parse HTML ---------------------------------------------------------

page <- read_html(file_path)

# Grab each listing "card".
# These selectors match the structure in your uploaded file.
# If Domain changes layout, try adjusting the selectors below.
cards <- html_elements(page, "ul.css-163kh4h > li div.css-qrqvvg")

# Fallback (if needed): try a more generic grab of containers with an H2 and a price-like <p>
if (length(cards) == 0) {
  cards <- html_elements(page, xpath = "//li[.//h2]")
}

extract_one <- function(card) {
  title_h2  <- card %>% html_element("h2")
  addr_raw  <- title_h2 %>% html_text2() %||% NA_character_

  # Price: first try the class seen in your file; then fall back to any <p> containing a $ or 'Offer'
  price_raw <- card %>% html_element("p.css-mgq8yx") %>% html_text2()
  if (is.null(price_raw)) {
    price_raw <- card %>% html_elements("p") %>% html_text2() %>% keep(~str_detect(.x, "\\$|OFFER|EOI|EXPRESSION|CONTACT|AUCTION")) %>% first()
  }
  price_raw <- price_raw %||% NA_character_

  tibble(Address_raw = addr_raw, Price_raw = price_raw)
}

raw_df <- map_dfr(cards, extract_one) %>% distinct()

# Keep rows that look like Tasmanian addresses
raw_df <- raw_df %>%
  filter(str_detect(Address_raw, "TAS\\s*\\d{4}") | str_detect(Address_raw, "TAS\\d{4}"))

# Split address into fields
addr_split <- map_dfr(raw_df$Address_raw, parse_address)

df <- bind_cols(raw_df, addr_split) %>%
  mutate(
    Price_numeric = map_dbl(Price_raw, parse_price_numeric),
    full_address  = str_squish(paste(street, suburb, state, postcode, "Australia"))
  )

# --- Geocoding ----------------------------------------------------------
# Option A: Free OpenStreetMap/Nominatim via tidygeocoder (no key, but slower; be courteous)
# Tip: For heavier use, consider setting your email in the query to be polite & improve reliability:
#      custom_query = list(email = "your_email@example.com")
#
# Note: min_time=1 spaces out requests to ~1/sec (respectful to OSM).
df_geo <- df %>%
  filter(!is.na(full_address)) %>%
  geocode(address = full_address,
          method = "osm",
          lat = latitude, long = longitude,
          limit = 1, min_time = 1,
          custom_query = list(countrycodes = "au"))

# Option B (faster/paid): Google Maps Geocoding
# library(ggmap)
# register_google(key = "YOUR_GOOGLE_API_KEY")
# df_geo <- df %>%
#   mutate_geocode(full_address, output = "latlon", source = "google") %>%
#   rename(latitude = lat, longitude = lon)

# Merge geocodes back (in case some rows failed to geocode)
df_final <- df %>%
  left_join(select(df_geo, full_address, latitude, longitude), by = "full_address")

# --- Output -------------------------------------------------------------
# See a preview
print(df_final, n = 50)

# Save to CSV
out_csv <- "./results/tas_listings_with_price_and_geocode.csv"
write_csv(df_final, out_csv)
message("Saved: ", out_csv)

```

## Determine Crop Suitability for Listig points
```{r}
# ======================== LIST ESM overlap for property points (robust) ========================
# Inputs:
#   CSV with lower-case columns: latitude, longitude  (plus any other reference columns)
# Outputs:
#   esm_suitability_by_property.csv  (long: property x crop)
#   esm_top_crops_by_property.csv    (filtered to Well/Suitable)
#   esm_wide_scores.csv              (wide matrix of scores per crop)
#
# install.packages(c("dplyr","readr","purrr","stringr","tibble","tidyr","httr2","jsonlite"))

library(dplyr)
library(readr)
library(purrr)
library(stringr)
library(tibble)
library(tidyr)
library(httr2)
library(jsonlite)

# ------------------------- CONFIG --------------------------------------

in_points_csv     <- "./results/tas_listings_with_price_and_geocode.csv"
out_full_results  <- "./results/esm_suitability_by_property.csv"
out_top_crops_csv <- "./results/esm_top_crops_by_property.csv"
out_wide_scores   <- "./results/esm_wide_scores.csv"

# ESM service (2018 baseline; swap to 2030/2050 if needed)
esm_base    <- "https://services.thelist.tas.gov.au/arcgis/rest/services/Public/EnterpriseSuitabilityMapping2018/MapServer"
identify_url<- paste0(esm_base, "/identify")
layers_url  <- paste0(esm_base, "/layers")

# Identify rendering params (tune if needed)
buf_m        <- 25000     # half-size of mapExtent (metres); larger helps scale/tolerance
tolerance_px <- 10        # pixel tolerance
img_w <- 1024; img_h <- 1024; img_dpi <- 96

# Try modes in this order per layer
identify_modes <- c("all", "visible", "top")

# Optional: sleep between Identify requests to be polite
identify_delay_sec <- 0.0

# ------------------------- HELPERS -------------------------------------

`%||%` <- function(a, b) if (is.null(a) || length(a) == 0) b else a
normalize_key <- function(x) gsub("[^a-z0-9]+", "", tolower(x))

# Vectorized bucketing
bucket_suitability <- function(val) {
  if (length(val) == 0) return(character(0))
  s <- as.character(val)
  out <- rep(NA_character_, length(s))
  num <- suppressWarnings(as.numeric(str_extract(s, "^\\d+(?:\\.\\d+)?")))
  idx_num <- !is.na(num)
  if (any(idx_num)) {
    major <- floor(num[idx_num])
    out[idx_num] <- dplyr::case_when(
      major == 1 ~ "Well suited",
      major == 2 ~ "Suitable",
      major == 3 ~ "Moderately suitable",
      major >= 4 ~ "Unsuitable",
      TRUE ~ NA_character_
    )
  }
  idx_txt <- which(is.na(out))
  if (length(idx_txt)) {
    su <- toupper(s[idx_txt])
    out[idx_txt[str_detect(su, "WELL")]]       <- "Well suited"
    out[idx_txt[str_detect(su, "UNSUIT")]]     <- "Unsuitable"
    out[idx_txt[str_detect(su, "MODERATE")]]   <- "Moderately suitable"
    mask_suit <- is.na(out[idx_txt]) & str_detect(su, "\\bSUIT")
    out[idx_txt[mask_suit]] <- "Suitable"
  }
  out
}

score_suitability <- function(group) {
  dplyr::case_when(
    group == "Well suited"         ~ 3,
    group == "Suitable"            ~ 2,
    group == "Moderately suitable" ~ 1,
    group == "Unsuitable"          ~ 0,
    TRUE                           ~ NA_real_
  )
}

# ---- 4326 -> 102100 (Web Mercator) reprojection ----
to_3857 <- function(lon, lat) {
  x <- lon * 20037508.34 / 180
  y <- log(tan((90 + lat) * pi / 360)) * 20037508.34 / pi
  list(x = x, y = y)
}

# -------- Robust layer discovery: handles data.frame OR list-of-lists ----
fetch_layers <- function() {
  resp <- request(layers_url) |> req_url_query(f = "json") |> req_perform()
  j <- resp_body_json(resp, simplifyVector = TRUE)

  lyr <- j$layers
  if (is.null(lyr)) stop("Service response has no 'layers' element.")

  if (is.data.frame(lyr)) {
    # Already a data frame: columns like id, name, type, minScale may be present
    df <- as_tibble(lyr) %>%
      transmute(
        id       = suppressWarnings(as.integer(id)),
        name     = as.character(name %||% NA_character_),
        type     = as.character(type %||% NA_character_),
        minScale = suppressWarnings(as.numeric(minScale %||% NA_real_))
      )
  } else if (is.list(lyr)) {
    # List of layer objects
    df <- map_dfr(lyr, function(x) tibble(
      id       = suppressWarnings(as.integer(x$id %||% NA_integer_)),
      name     = as.character(x$name %||% NA_character_),
      type     = as.character(x$type %||% NA_character_),
      minScale = suppressWarnings(as.numeric(x$minScale %||% NA_real_))
    ))
  } else {
    stop("Unexpected 'layers' type: ", typeof(lyr))
  }

  df %>%
    filter(!is.na(id)) %>%
    arrange(id) %>%
    # keep the raster enterprise layers
    filter(grepl("raster", type, ignore.case = TRUE))
}

# Identify ONE layer at a point, trying modes 'all:', 'visible:', 'top:'
identify_point_one_layer <- function(lon4326, lat4326, layer_id) {
  wm <- to_3857(lon4326, lat4326)
  xmin <- wm$x - buf_m; ymin <- wm$y - buf_m
  xmax <- wm$x + buf_m; ymax <- wm$y + buf_m

  for (mode in identify_modes) {
    req <- request(identify_url) |>
      req_url_query(
        f = "json",
        geometry = sprintf('{"x":%.3f,"y":%.3f,"spatialReference":{"wkid":102100}}', wm$x, wm$y),
        geometryType = "esriGeometryPoint",
        sr = 102100,
        layers = paste0(mode, ":", layer_id),
        tolerance = tolerance_px,
        mapExtent = paste(xmin, ymin, xmax, ymax, sep = ","),
        imageDisplay = paste(img_w, img_h, img_dpi, sep = ","),
        returnGeometry = "false",
        returnUnformattedValues = "true",
        returnFieldName = "true"
      )

    resp <- tryCatch(req_perform(req), error = function(e) NULL)
    if (is.null(resp)) next

    out <- tryCatch(resp_body_json(resp, simplifyVector = TRUE), error = function(e) NULL)
    if (is.null(out) || is.null(out$results) || length(out$results) == 0) next

    r <- tibble::as_tibble(out$results)
    if (nrow(r) == 0) next
    return(r)
  }
  tibble() # no hit
}

# Extract class value (pixel) from Identify results
pluck_identify_values <- function(r) {
  n <- nrow(r)
  if (n == 0) return(character(0))

  # Case A: top-level 'value'
  if ("value" %in% names(r) && !all(is.na(r$value))) {
    return(as.character(r$value))
  }

  vals <- rep(NA_character_, n)
  if ("attributes" %in% names(r)) {
    wanted <- c("pixelvalue","pixel_value","rastervalue","value",
                "class","classcode","suit","suit_class","suitability")
    for (i in seq_len(n)) {
      att <- r$attributes[[i]]
      if (is.null(att)) next

      # scalar
      if (is.atomic(att) && length(att) == 1 && is.null(names(att))) {
        vals[i] <- as.character(att); next
      }
      # atomic named vector
      if (is.atomic(att) && length(att) >= 1 && !is.null(names(att))) {
        nms <- names(att); norm <- normalize_key(nms)
        hit <- which(norm %in% wanted)
        vals[i] <- if (length(hit)) as.character(att[[hit[1]]]) else as.character(att[[1]])
        next
      }
      # list
      if (is.list(att)) {
        nms <- names(att)
        if (!is.null(nms)) {
          norm <- normalize_key(nms)
          hit <- which(norm %in% wanted)
          if (length(hit)) {
            vals[i] <- as.character(att[[hit[1]]])
          } else {
            flat <- unlist(att, recursive = TRUE, use.names = FALSE)
            if (length(flat)) vals[i] <- as.character(flat[1])
          }
        } else {
          flat <- unlist(att, recursive = TRUE, use.names = FALSE)
          if (length(flat)) vals[i] <- as.character(flat[1])
        }
        next
      }
    }
  }
  vals
}

# ------------------------- LOAD POINTS ---------------------------------

pts_in <- read_csv(in_points_csv, show_col_types = FALSE)

if (!all(c("latitude","longitude") %in% names(pts_in))) {
  stop("Input CSV must have lower-case 'latitude' and 'longitude' columns.")
}

if (!"prop_id" %in% names(pts_in)) {
  pts_in <- pts_in %>% mutate(prop_id = row_number())
} else {
  pts_in <- pts_in %>% mutate(prop_id = as.integer(factor(prop_id)))
}

pts <- pts_in %>%
  filter(!is.na(latitude), !is.na(longitude)) %>%
  select(prop_id, everything())

message("Properties to query: ", nrow(pts))

# ------------------------- DISCOVER LAYERS -----------------------------

layers_df <- fetch_layers()
stopifnot(nrow(layers_df) > 0)

layer_ids   <- layers_df$id
layer_names <- setNames(layers_df$name, as.character(layers_df$id))

message("Layers discovered: ",
        paste0(layer_ids, ":", layer_names[as.character(layer_ids)], collapse = "; "))

# ------------------------- QUERY ESM -----------------------------------

results <- pts %>%
  split(.$prop_id) %>%
  map_dfr(function(df1) {
    lon <- df1$longitude[1]; lat <- df1$latitude[1]

    per_layer <- map(layer_ids, function(lyr_id) {
      r <- identify_point_one_layer(lon, lat, lyr_id)
      if (identify_delay_sec > 0) Sys.sleep(identify_delay_sec)
      if (nrow(r) == 0) return(tibble())
      vals <- pluck_identify_values(r)
      tibble(
        prop_id    = df1$prop_id[1],
        layerId    = lyr_id,
        layerName  = layer_names[as.character(lyr_id)] %||%
                     if ("layerName" %in% names(r)) r$layerName[1] else NA_character_,
        class_code = vals
      )
    })

    bind_rows(per_layer)
  })

if (nrow(results) == 0) {
  stop("Identify returned no results. Increase buf_m, check network/firewall, or try the 2030/2050 service URLs.")
}

final <- results %>%
  mutate(
    suitability_group = bucket_suitability(class_code),
    suitability_score = score_suitability(suitability_group),
    crop              = make.names(layerName)
  ) %>%
  select(prop_id, crop, class_code, suitability_group, suitability_score) %>%
  left_join(pts_in %>% select(prop_id, any_of(c(
    "street","suburb","state","postcode","price_raw","price_numeric",
    "address_raw","full_address","latitude","longitude"
  ))), by = "prop_id") %>%
  arrange(prop_id, crop)

write_csv(final, out_full_results)
message("Saved: ", out_full_results)

top <- final %>%
  filter(suitability_group %in% c("Well suited","Suitable")) %>%
  group_by(prop_id, across(all_of(intersect(names(final), c(
    "street","suburb","state","postcode","price_raw","price_numeric",
    "address_raw","full_address","latitude","longitude"
  ))))) %>%
  summarise(
    crops = paste0(crop, " (", suitability_group, ":", class_code, ")", collapse = "; "),
    best_score = max(suitability_score, na.rm = TRUE),
    n_crops    = n(),
    .groups = "drop"
  )

write_csv(top, out_top_crops_csv)
message("Saved: ", out_top_crops_csv)

wide_scores <- final %>%
  select(prop_id, crop, suitability_score) %>%
  distinct() %>%
  pivot_wider(names_from = crop, values_from = suitability_score)

write_csv(wide_scores, out_wide_scores)
message("Saved: ", out_wide_scores)

message("Done.")

```

## Function for determinnig crop suitability at any point

```{r}
# --- Packages ---
library(httr2)
library(jsonlite)
library(dplyr)
library(purrr)
library(tibble)
library(stringr)

# --- Service chooser: "2018" (baseline) or "2030" (CFT RCP8.5) ---
esm_base_url <- function(service_year = c("2018","2030")) {
  service_year <- match.arg(service_year)
  sprintf(
    "https://services.thelist.tas.gov.au/arcgis/rest/services/Public/EnterpriseSuitabilityMapping%s/MapServer",
    service_year
  )
}

# --- Helpers ---
json_get <- function(url, query = list()) {
  # LIST often serves JSON as text/plain; allow it.
  request(url) |>
    req_url_query(!!!query) |>
    req_perform() |>
    resp_body_json(check_type = FALSE, simplifyVector = FALSE)
}

# WGS84 lon/lat -> Web Mercator (EPSG:3857)
lonlat_to_mercator <- function(lon, lat) {
  x <- lon * 20037508.34 / 180
  y <- log(tan((90 + lat) * pi/360)) * 20037508.34 / pi
  list(x = x, y = y)
}

# Build readable, unique column names from layer names (preserve bracket tags)
make_colnames <- function(layer_names, layer_ids) {
  keep_brackets <- str_replace_all(layer_names, "\\[\\s*([^\\]]+?)\\s*\\]", "_\\1")
  base <- keep_brackets |>
    tolower() |>
    str_replace_all("[^a-z0-9]+", "_") |>
    str_replace_all("_+", "_") |>
    str_replace_all("^_|_$", "")
  unique_names <- make.unique(base, sep = "_")
  if (any(duplicated(unique_names))) {
    dup <- duplicated(unique_names) | duplicated(unique_names, fromLast = TRUE)
    unique_names[dup] <- paste0(unique_names[dup], "_", layer_ids[dup])
  }
  unique_names
}

# --- Get crop layers (exclude versatility index overlays) ---
get_crop_layers <- function(base_url) {
  svc <- json_get(base_url, query = list(f = "pjson"))
  layers <- svc$layers
  tibble(
    id   = vapply(layers, function(z) z$id, integer(1)),
    name = vapply(layers, function(z) z$name, character(1))
  ) |>
    filter(!str_detect(tolower(name), "versatility")) |>
    arrange(id)
}

# --- Legend (layerId -> named vector code->label); used only as fallback ---
get_legend_maps <- function(base_url) {
  leg <- json_get(paste0(base_url, "/legend"), query = list(f = "pjson"))
  out <- map(leg$layers, function(L) {
    if (is.null(L$legend)) return(NULL)
    labs <- vapply(L$legend, function(v) if (is.null(v$label)) "" else as.character(v$label), character(1))
    labs <- str_squish(labs)
    code <- str_match(labs, "^([0-9]+(?:\\.[0-9]+)?)")[,2]
    keep <- !is.na(code)
    if (!any(keep)) return(NULL)
    stats::setNames(labs[keep], code[keep])
  })
  names(out) <- vapply(leg$layers, function(L) as.character(L$layerId), character(1))
  out
}

# --- Convert raw numeric to codes like "1.0","1.1" ... (used in fallback) ---
normalize_code <- function(x) {
  if (is.na(x) || x == "") return(NA_character_)
  x <- str_squish(as.character(x))
  # If a code appears in the text, grab it
  m <- str_match(x, "([0-9]+(?:\\.[0-9]+)?)")
  if (!is.na(m[1,2])) return(m[1,2])
  # Otherwise try numeric divisions
  xn <- suppressWarnings(as.numeric(x))
  if (is.na(xn)) return(NA_character_)
  for (div in c(1,10,100,1000)) {
    v <- xn/div
    code <- sprintf("%.1f", v)
    if (str_detect(code, "^[0-9]+\\.[0-9]$")) return(code)
  }
  NA_character_
}

# --- Identify and extract both Suitability and a raw value (for fallback) ---
identify_suitability_and_raw <- function(base_url, lon, lat, layer_ids) {
  pt <- lonlat_to_mercator(lon, lat)
  buf <- 1000
  map_extent <- sprintf("%.3f,%.3f,%.3f,%.3f", pt$x - buf, pt$y - buf, pt$x + buf, pt$y + buf)

  q <- list(
    f = "json",
    geometry = jsonlite::toJSON(list(x = pt$x, y = pt$y), auto_unbox = TRUE),
    geometryType = "esriGeometryPoint",
    sr = 3857,
    layers = paste0("all:", paste(layer_ids, collapse = ",")),
    tolerance = 1,
    mapExtent = map_extent,
    imageDisplay = "800,600,96",
    returnGeometry = "false",
    returnUnformattedValues = "true",
    returnFieldName = "false"
  )

  res <- json_get(paste0(base_url, "/identify"), query = q)

  suit <- rep(NA_character_, length(layer_ids)); names(suit) <- as.character(layer_ids)
  raw  <- rep(NA_character_, length(layer_ids)); names(raw)  <- as.character(layer_ids)

  if (!is.null(res$results) && length(res$results)) {
    for (r in res$results) {
      lid <- as.character(r$layerId)
      if (!is.null(r$attributes) && length(r$attributes)) {
        # 1) Pull Suitability (exact field) if present (case/space tolerant)
        keys <- names(r$attributes)
        normkey <- function(k) gsub("\\s+", "", tolower(k))
        idx_suit <- which(normkey(keys) == "suitability")
        if (length(idx_suit)) {
          v <- r$attributes[[ idx_suit[1] ]]
          if (length(v) == 1) suit[lid] <- as.character(v)
        }
        # 2) Keep a raw-ish numeric for legend fallback
        if (is.na(raw[lid])) {
          # look for a value-like field
          hit <- which(str_detect(tolower(keys), "(pixel|value|class|classcode|rastervalue|suitcode)"))
          if (length(hit)) {
            for (h in hit) {
              v <- r$attributes[[h]]
              if (length(v) == 1) { raw[lid] <- as.character(v); break }
            }
          } else {
            first_scalar <- r$attributes[[1]]
            if (length(first_scalar) == 1) raw[lid] <- as.character(first_scalar)
          }
        }
      }
    }
  }
  list(suitability = suit, raw = raw)
}

# --- If Suitability missing, use legend to map raw code -> label ---
decode_with_legend <- function(raw_by_layer, legend_maps) {
  out <- rep(NA_character_, length(raw_by_layer)); names(out) <- names(raw_by_layer)
  for (lid in names(raw_by_layer)) {
    code <- normalize_code(raw_by_layer[[lid]])
    lm <- legend_maps[[lid]]
    if (!is.null(lm) && !is.na(code) && code %in% names(lm)) {
      out[[lid]] <- lm[[code]]
    } else {
      out[[lid]] <- code
    }
  }
  out
}

# --- Main: one-row tibble with lon, lat, and Suitability per crop layer ---
esm_point_suitability <- function(lon, lat, service_year = c("2018","2030")) {
  base_url   <- esm_base_url(match.arg(service_year))
  layers_tbl <- get_crop_layers(base_url)
  layer_ids  <- layers_tbl$id

  # Identify once
  id <- identify_suitability_and_raw(base_url, lon, lat, layer_ids)

  # Fallback legend decode only where Suitability is missing
  need_fallback <- which(is.na(id$suitability) | id$suitability == "")
  label_fallback <- character(0)
  if (length(need_fallback)) {
    legend_maps <- get_legend_maps(base_url)
    label_fallback <- decode_with_legend(id$raw[need_fallback], legend_maps)
  }

  labels <- id$suitability
  if (length(need_fallback)) labels[names(label_fallback)] <- label_fallback

  # Order & name columns
  labels <- labels[as.character(layer_ids)]
  col_names <- make_colnames(layers_tbl$name, layer_ids = as.character(layer_ids))
  names(labels) <- col_names

  tibble(
    lon = lon,
    lat = lat,
    !!!as.list(labels)
  )
}

# ---------- Example: your test coordinate ----------
lon <- 147.8070861
lat <- -39.9014742

lon <- 147.9450
lat <- -42.32671

# Baseline 2018 ESM; values come from the 'Suitability' attribute where available
df_2018 <- esm_point_suitability(lon, lat, service_year = "2018")
print(df_2018)

# If you also want the 2030 projection (CFT RCP8.5), use:
# df_2030 <- esm_point_suitability(lon, lat, service_year = "2030")
# print(df_2030)


```


## Check the layer and point location
```{r}
# install.packages(c("leaflet"), dependencies = TRUE)  # if needed
library(leaflet)

esm_wms_base <- function(service_year = c("2018","2030","2050")) {
  service_year <- match.arg(service_year)
  sprintf("https://services.thelist.tas.gov.au/arcgis/services/Public/EnterpriseSuitabilityMapping%s/MapServer/WMSServer",
          service_year)
}

# Quick interactive preview of ONE crop layer at your point
esm_leaflet_preview <- function(lon, lat, layer_id = 0, service_year = c("2018","2030"), zoom = 12, opacity = 0.75) {
  service_year <- match.arg(service_year)
  base <- esm_wms_base(service_year)
  leaflet() |>
    addProviderTiles("CartoDB.Positron") |>
    addWMSTiles(
      baseUrl = base,
      layers = as.character(layer_id),   # e.g. 0=Barley, 19=Potatoes, 24=Sparkling wine grapes
      options = WMSTileOptions(format = "image/png", transparent = TRUE),
      group = paste0("ESM ", service_year, " layer ", layer_id)
    ) |>
    addCircleMarkers(lng = lon, lat = lat, radius = 6, weight = 2, fillOpacity = 0.9,
                     popup = paste0("Lon: ", lon, "<br>Lat: ", lat)) |>
    setView(lng = lon, lat = lat, zoom = zoom) |>
    addLayersControl(overlayGroups = paste0("ESM ", service_year, " layer ", layer_id), options = layersControlOptions(collapsed = FALSE))
}

# Example at your point — try a few layer IDs:
#lon <- 147.8070861
#lat <- -39.9014742

lon <- 147.9450
lat <- -42.32671


#esm_leaflet_preview(lon, lat, layer_id = 0,  service_year = "2018")  # Barley (0)
# esm_leaflet_preview(lon, lat, layer_id = 19, service_year = "2018") # Potatoes (19)
# esm_leaflet_preview(lon, lat, layer_id = 24, service_year = "2018") # Sparkling Wine Grapes (24)
```


## Issues I have identified

Using a geolocated single point - sometimes falls on a roadway (e.g the post box of the address) - making the entire property 'unsuitable' for all crops. What we need to do is get the actual parcel/polygon from the cadaster for the adress and then use this to find the majority pixel value inside the parcel for each crop 

## Using Cadaster PID

In this section I give an example of how to query the ListMap WMS address layer to get the property ID (PID) -> This links to the cadaster polygon. Then we compute (by majority pixel) the crop suitability inside the cadaster polygon.

### Function for returning ESM majorities by parcel

**Purpose**  
Resolve a street address to its cadastral parcel (PID) using LIST web services, then compute **majority crop suitability** for that parcel from the Enterprise Suitability Mapping (ESM) **without downloading rasters**.

**How it works**
1. **Address → PID** via LIST Address Points (tolerant to minor typos).
2. **PID → parcel polygon** from LIST Cadastral Parcels (projected to EPSG:3857).
3. **Crop layers**: if `crop_names = NULL`, queries **all** leaf ESM raster layers; otherwise resolves partial, case-insensitive names (e.g., `"potato"`).
4. **Sampling**: builds a point grid inside the parcel (`cellsize_m` metres), calls the MapServer **Identify** at each point, reads numeric class **and** text **Suitability** (falls back to the layer legend if needed).
5. **Aggregation**: tallies samples per (code, label) and reports the **majority class** per crop.

**Signature**
```{r}
esm_majority_for_address(
  address_text,
  crop_names = NULL,   # NULL = all crops
  cellsize_m = 60,     # grid spacing in metres (speed/accuracy trade-off)
  max_pts   = 2000     # cap on identify requests
)
```

**Returns** (named list)
- `pid` — matched LIST PID  
- `address` — human-readable address string  
- `address_record` — LIST Address Points feature (`sf`)  
- `parcel` — parcel polygon (`sf`, EPSG:3857)  
- `crops` — tibble mapping `layer_id ↔ crop`  
- `summary` — tibble with:
  - `address`, `pid`, `crop`, `layer_id`, `code`, `label`, `pct`, `n`, `rank`  
  - `rank == 1` marks the majority class for that crop

**Typical use**
```{r}
out <- esm_majority_for_address(
  "850 West End Rd, Leeka 7255, TAS",
  crop_names = NULL,     # all crops
  cellsize_m = 60,
  max_pts = 1200
)

dplyr::filter(out$summary, rank == 1) |>
  dplyr::select(crop, code, label, pct, n, address, pid) |>
  dplyr::arrange(crop)
```

**Notes**
- Results are **sample-count majorities** (proxy for area). Exact area-weighted classes would require server-side zonal stats or local rasters.  
- Keep `cellsize_m` reasonable (e.g., 60–100 m) and `max_pts` capped to avoid hammering the service.  
- Labels come from each layer’s `Suitability` attribute when present; the legend is used as a fallback.

### Actual code

```{r}
# ================== LIST REST → PID → Parcel → ESM majority (by crop name) ==================
# Requirements: sf, dplyr, stringr, httr2, jsonlite, purrr, tidyr

library(sf)
library(dplyr)
library(stringr)
library(httr2)
library(jsonlite)
library(purrr)
library(tidyr)

# --- Endpoints (LIST Tas) -----------------------------------------------------
addr_layer <- "https://services.thelist.tas.gov.au/arcgis/rest/services/Public/OpenDataWFS/MapServer/9"
parc_layer <- "https://services.thelist.tas.gov.au/arcgis/rest/services/Public/OpenDataWFS/MapServer/14"
esm_ms     <- "https://services.thelist.tas.gov.au/arcgis/rest/services/Public/EnterpriseSuitabilityMapping2018/MapServer"
sr_epsg    <- 3857

# --- Helper: safe URL builder (handles vector params) -------------------------
url_query <- function(base, params) {
  kv <- vapply(names(params), function(k){
    val <- params[[k]]
    if (length(val) > 1) val <- paste(val, collapse = ",")
    paste0(k, "=", URLencode(as.character(val), reserved = TRUE))
  }, "")
  paste0(base, "?", paste(kv, collapse = "&"))
}

# --- Helper: robust JSON fetcher (tolerates text/plain JSON) ------------------
get_json <- function(url) {
  resp <- request(url) |>
    req_headers("Accept" = "application/json, text/plain, */*") |>
    req_perform()
  txt <- resp_body_string(resp)
  jsonlite::fromJSON(txt, simplifyVector = FALSE)
}

# --- 1) Address → PID (tolerant to "Lekka" vs "Leeka") ------------------------
address_to_pid <- function(address_text){
  x  <- toupper(address_text)
  pc <- stringr::str_extract(x, "\\b\\d{4}\\b")
  hn <- suppressWarnings(as.integer(stringr::str_extract(x, "\\b\\d+\\b")))
  street_all <- stringr::str_match(x, "\\b\\d+\\s+(.+?),\\s*[A-Z ]+\\s*\\d{4}")[,2] |> stringr::str_squish()
  if (is.na(street_all)) stop("Could not parse street from address: ", address_text)

  tokens <- unlist(strsplit(street_all, "\\s+"))
  st_type_guess <- tail(tokens, 1)
  street_guess  <- stringr::str_squish(gsub(paste0("\\s+", st_type_guess, "$"), "", street_all))

  base <- paste0(addr_layer, "/query")

  # First pass: number + postcode + street LIKE
  q1 <- url_query(base, list(
    where = paste0("ST_NO_FROM=", hn,
                   " AND POSTCODE=", pc,
                   " AND UPPER(STREET) LIKE '", street_guess, "%'"),
    outFields="*", returnGeometry="true", outSR=sr_epsg, f="geojson"
  ))
  ap <- suppressWarnings(sf::st_read(q1, quiet = TRUE))

  if (nrow(ap) == 0) {
    # Fallback: drop house number, pick closest by number afterwards
    q2 <- url_query(base, list(
      where = paste0("POSTCODE=", pc,
                     " AND UPPER(STREET) LIKE '", street_guess, "%'"),
      outFields="*", returnGeometry="true", outSR=sr_epsg, f="geojson"
    ))
    ap <- suppressWarnings(sf::st_read(q2, quiet = TRUE))
    if (nrow(ap) == 0) stop("No LIST address match for: ", address_text)
    ap$NO <- suppressWarnings(as.integer(ap$ST_NO_FROM))
    ap <- ap[order(abs(ap$NO - hn)), ][1, ]
  } else {
    ap <- ap[1, ]
  }

  if (!"PID" %in% names(ap)) stop("LIST Address Points response lacks PID field.")
  ap
}

# Build a tidy, human-readable address string from the Address Points record
pretty_address <- function(ap_row){
  if ("FULL_ADDRESS" %in% names(ap_row) && !is.na(ap_row$FULL_ADDRESS[1])) {
    return(as.character(ap_row$FULL_ADDRESS[1]))
  }
  parts <- c(
    ap_row$ST_NO_FROM, ap_row$STREET, ap_row$ST_TYPE,
    ap_row$LOCALITY, ap_row$POSTCODE
  )
  paste(na.omit(as.character(parts)), collapse = " ")
}

# --- 2) PID → Parcel polygon --------------------------------------------------
parcel_by_pid <- function(pid){
  q <- url_query(paste0(parc_layer,"/query"), list(
    where = paste0("PID=", pid),
    outFields="PID,COMP_AREA,CAD_TYPE1,TENURE_TY",
    returnGeometry="true", outSR=sr_epsg, f="geojson"
  ))
  p <- suppressWarnings(sf::st_read(q, quiet = TRUE))
  if (nrow(p) == 0) stop("No parcel found for PID: ", pid)
  p[1,]
}

# --- 3) Discover crop layers (id ↔ name) -------------------------------------
esm_layers <- function(){
  lyr <- get_json(url_query(paste0(esm_ms, "/layers"), list(f="pjson")))
  layers <- lyr$layers
  tibble(
    layer_id = vapply(layers, function(x) x$id, 0),
    crop     = vapply(layers, function(x) x$name, ""),
    type     = vapply(layers, function(x) x$type %||% "", "")
  ) |>
    # keep leaf raster layers (ignore group/annotation/etc.)
    filter(!grepl("Group", type, ignore.case = TRUE)) |>
    filter(grepl("Raster|Mosaic", type, ignore.case = TRUE))
}

# Resolve user-provided crop names (partial, case-insensitive) to layer ids
resolve_crops <- function(crop_names = NULL){
  tbl <- esm_layers()
  if (is.null(crop_names) || length(crop_names) == 0) return(tbl)  # all crops
  pats <- paste0("(", paste0(stringr::str_replace_all(crop_names, "[^A-Za-z0-9]+", ".*"), collapse = "|"), ")")
  tbl |> filter(grepl(pats, crop, ignore.case = TRUE))
}

# --- 4) Legend lookup (code → label) for an ESM layer ------------------------
esm_legend_lookup <- function(layer_id){
  leg <- get_json(url_query(paste0(esm_ms, "/legend"), list(f="pjson")))
  ids <- vapply(leg$layers, function(x) x$layerId, 0)
  idx <- which(ids == layer_id)
  if (!length(idx)) return(tibble(code = numeric(), label = character()))
  lyr <- leg$layers[[idx]]
  labs <- vapply(lyr$legend, `[[`, "", "label")
  codes <- stringr::str_match(labs, "^(\\d+(?:\\.\\d+)?)")[,2]
  names <- stringr::str_match(labs, "^(?:\\d+(?:\\.\\d+)?)\\s+(.+)$")[,2]
  tibble(code = as.numeric(codes), label = names) |> distinct()
}

# --- Helpers to pick code/label from Identify attributes ----------------------
`%||%` <- function(a, b) if (!is.null(a)) a else b

.pick_numeric_code <- function(atts){
  nums <- atts[vapply(atts, is.numeric, TRUE)]
  if (!length(nums)) return(NA_real_)
  nms <- names(nums)
  pref <- grepl("(?i)(code|class|value|suit)", nms)
  idx <- if (any(pref)) which(pref)[1] else 1
  as.numeric(nums[[idx]])
}
.pick_label <- function(atts){
  chars <- atts[vapply(atts, is.character, TRUE)]
  if (!length(chars)) return(NA_character_)
  nms <- names(chars)
  # prefer "Suitability" or anything containing 'suit'
  exact <- which(grepl("(?i)^suit(ability)?$", nms))
  if (length(exact)) return(as.character(chars[[ exact[1] ]]))
  soft  <- which(grepl("(?i)suit", nms))
  if (length(soft))  return(as.character(chars[[ soft[1] ]]))
  # fallback: first longest-looking string value
  as.character(chars[[ which.max(nchar(unlist(chars))) ]])
}

# --- 5) Identify: return BOTH code and label for one point --------------------
esm_identify_point <- function(xy_3857, layer_id, map_extent_3857, img_px = c(1000,1000,96)){
  id_url <- paste0(esm_ms, "/identify")
  geom_json <- jsonlite::toJSON(
    list(x = xy_3857[1], y = xy_3857[2], spatialReference = list(wkid=sr_epsg)),
    auto_unbox = TRUE
  )
  params <- list(
    f="json",
    geometry = geom_json,
    geometryType = "esriGeometryPoint",
    sr = as.character(sr_epsg),
    layers = paste0("all:", layer_id),
    tolerance = 1,
    mapExtent = paste(map_extent_3857, collapse=","),
    imageDisplay = paste(img_px, collapse=","),
    returnGeometry = "false"
  )
  r <- get_json(url_query(id_url, params))
  if (length(r$results)==0) return(tibble(code=NA_real_, label=NA_character_))
  atts <- r$results[[1]]$attributes

  code  <- .pick_numeric_code(atts)
  label <- .pick_label(atts)

  if (!is.finite(code) && is.character(label)) {
    code <- suppressWarnings(as.numeric(stringr::str_match(label, "^(\\d+(?:\\.\\d+)?)")[,2]))
  }

  tibble(code = as.numeric(code), label = ifelse(is.na(label), NA_character_, as.character(label)))
}

# --- 6) Sample grid → identify → aggregate by (code,label) --------------------
parcel_majority_for_layer <- function(parcel_3857, layer_id, crop_name, cellsize_m = 60, max_pts = 2000){
  bbox <- sf::st_bbox(parcel_3857)
  map_extent <- c(bbox["xmin"], bbox["ymin"], bbox["xmax"], bbox["ymax"])

  grid <- sf::st_make_grid(parcel_3857, cellsize = cellsize_m, what = "centers", square = TRUE)
  pts  <- sf::st_as_sf(grid, crs = sr_epsg) |> suppressMessages(sf::st_filter(parcel_3857))
  if (nrow(pts) == 0) {
    return(tibble(crop=crop_name, layer_id=layer_id, code=NA_real_, label=NA_character_, pct=NA_real_, n=0))
  }
  if (nrow(pts) > max_pts) {
    set.seed(1)
    pts <- pts[sample.int(nrow(pts), max_pts), , drop = FALSE]
  }
  xy  <- sf::st_coordinates(pts)

  vals <- map(seq_len(nrow(xy)), ~ esm_identify_point(xy[.x,], layer_id, map_extent))
  df   <- dplyr::bind_rows(vals) |> mutate(code = round(as.numeric(code), 1))

  # Fill missing labels using the layer legend
  if (any(is.na(df$label))) {
    lut <- esm_legend_lookup(layer_id)
    if (nrow(lut)) {
      df <- df |> left_join(lut, by = "code", suffix = c("", "_lut")) |>
        mutate(label = coalesce(label, label_lut)) |>
        select(-label_lut)
    }
  }

  tab <- df |>
    filter(is.finite(code) | !is.na(label)) |>
    count(code, label, name = "n") |>
    mutate(pct = n / sum(n)) |>
    arrange(desc(pct)) |>
    mutate(crop = crop_name, layer_id = layer_id, .before = 1)

  if (!nrow(tab)) tab <- tibble(crop=crop_name, layer_id=layer_id, code=NA_real_, label=NA_character_, pct=NA_real_, n=0)
  tab
}

# --- 7) Wrapper: Address → PID → Parcel → Majority per crop -------------------
esm_majority_for_address <- function(address_text,
                                     crop_names = NULL,      # NULL = all crops
                                     cellsize_m = 60, max_pts = 2000) {
  adr  <- address_to_pid(address_text)
  pid  <- adr$PID[1]
  addr_str <- pretty_address(adr)

  parc <- parcel_by_pid(pid) |> sf::st_transform(sr_epsg)

  # Resolve crops to layer_ids
  crops_tbl <- resolve_crops(crop_names)
  if (!nrow(crops_tbl)) stop("No ESM layers matched your crop names.")

  res_list <- purrr::pmap(
    list(crops_tbl$layer_id, crops_tbl$crop),
    ~ parcel_majority_for_layer(parc, layer_id = ..1, crop_name = ..2, cellsize_m = cellsize_m, max_pts = max_pts)
  )
  res <- dplyr::bind_rows(res_list) |>
    group_by(crop) |>
    mutate(rank = row_number(desc(pct))) |>
    ungroup() |>
    mutate(address = addr_str, pid = pid, .before = 1)

  list(pid = pid, address = addr_str, address_record = adr, parcel = parc, summary = res, crops = crops_tbl)
}

# ============================= Example runs ===================================

# 1) Default: ALL crops from the service
# out_all <- esm_majority_for_address(
#   "850 West End Rd, Lekka 7255, TAS",
#   crop_names = NULL,           # NULL → all crops
#   cellsize_m = 60,
#   max_pts = 1200
# )

# Inspect majority per crop (human-readable)
# out_all$summary %>%
#   filter(rank == 1) %>%
#   select(crop, code, label, pct, n, address, pid) %>%
#   arrange(crop)

# 2) Or choose specific crops by name (partial matches ok)
out_some <- esm_majority_for_address(
  "401 Bresnehans Road, LITTLE SWANPORT TAS 7190",
  crop_names = c("barley", "potato", "sparkling wine"),
  cellsize_m = 60,
  max_pts = 1200
)

out_some$summary %>%
  filter(rank == 1) %>%
  select(crop, code, label, pct, n, address, pid)

plot(out_some$parcel["PID"])

```

