---
title: "Data_pipeline"
format: html
editor: visual
---

```{r}
library(DBI)
library(RSQLite)
library(tidyverse)
library(dplyr)
library(dbplyr)
library(lubridate)
library(sf)
library(duckdb)
```

-   Reading the cadstral data for each municipality in Tasmania

```{r}
parcels_break_o_day <- st_read("dataset/LIST_PARCELS_BREAK_O_DAY/list_parcels_break_o_day.shp")
parcels_brighton<- st_read("dataset/LIST_PARCELS_BRIGHTON/list_parcels_brighton.shp")
parcels_burnie <- st_read("dataset/LIST_PARCELS_BURNIE/list_parcels_burnie.shp")
parcels_central_coast <- st_read("dataset/LIST_PARCELS_CENTRAL_COAST/list_parcels_central_coast.shp")
parcels_central_highlands <- st_read("dataset/LIST_PARCELS_CENTRAL_HIGHLANDS/list_parcels_central_highlands.shp")
parcels_circular_head <- st_read("dataset/LIST_PARCELS_CIRCULAR_HEAD/list_parcels_circular_head.shp")
parcels_clarence <- st_read("dataset/LIST_PARCELS_CLARENCE/list_parcels_CLARENCE.shp")
parcels_derwent_valley <- st_read("dataset/LIST_PARCELS_DERWENT_VALLEY/list_parcels_derwent_valley.shp")
parcels_devonport <- st_read("dataset/LIST_PARCELS_DEVONPORT/list_parcels_devonport.shp")
parcels_dorset<- st_read("dataset/LIST_PARCELS_DORSET/list_parcels_dorset.shp")
parcels_flinders<- st_read("dataset/LIST_PARCELS_FLINDERS/list_parcels_flinders.shp")
parcels_george_town <- st_read("dataset/LIST_PARCELS_GEORGE_TOWN/list_parcels_george_town.shp")
parcels_glamorgan_spring_bay <- st_read("dataset/LIST_PARCELS_GLAMORGAN_SPRING_BAY/list_parcels_glamorgan_spring_bay.shp")
parcels_glenorchy <- st_read("dataset/LIST_PARCELS_glenorchy/list_parcels_glenorchy.shp")
parcels_hobart <- st_read("dataset/LIST_PARCELS_HOBART/list_parcels_hobart.shp")
parcels_huon_valley <- st_read("dataset/LIST_PARCELS_HUON_VALLEY/list_parcels_huon_valley.shp")
parcels_kentish <- st_read("dataset/LIST_PARCELS_KENTISH/list_parcels_kentish.shp")
parcels_kingborough <- st_read("dataset/LIST_PARCELS_KINGBOROUGH/list_parcels_kingborough.shp")
parcels_king_island <- st_read("dataset/LIST_PARCELS_KING_ISLAND/list_parcels_king_island.shp")
parcels_latrobe <- st_read("dataset/LIST_PARCELS_LATROBE/list_parcels_latrobe.shp")
parcels_launceston <- st_read("dataset/LIST_PARCELS_LAUNCESTON/list_parcels_launceston.shp")
parcels_meander_valley <- st_read("dataset/LIST_PARCELS_MEANDER_VALLEY/list_parcels_meander_valley.shp")
parcels_northern_midlands <- st_read("dataset/LIST_PARCELS_NORTHERN_MIDLANDS/list_parcels_northern_midlands.shp")
parcels_sorell <- st_read("dataset/LIST_PARCELS_SORELL/list_parcels_sorell.shp")
parcels_southern_midlands <- st_read("dataset/LIST_PARCELS_SOUTHERN_MIDLANDS/list_parcels_southern_midlands.shp")
parcels_tasman <- st_read("dataset/LIST_PARCELS_TASMAN/list_parcels_tasman.shp")
parcels_waratah_wynyard <- st_read("dataset/LIST_PARCELS_WARATAH_WYNYARD/list_parcels_waratah_wynyard.shp")
parcels_west_coast <- st_read("dataset/LIST_PARCELS_WEST_COAST/list_parcels_west_coast.shp")
parcels_west_tamar <- st_read("dataset/LIST_PARCELS_WEST_TAMAR/list_parcels_west_tamar.shp")
```

-   Reading the Address Master data and property listing datax

```{r}
Address_Master <- st_read("dataset/LIST_ADDRESS_POINTS_STATEWIDE/list_address_points_statewide.shp")
view(Address_Master)
property_listing_domain<- read_csv("dataset/tas_listings_with_price_and_geocode.csv")
view(property_listing_domain)

Address_Master$ST_NO_FROM <- as.character(Address_Master$ST_NO_FROM)
Address_Master$POSTCODE <- as.character(Address_Master$POSTCODE)
```

-   Creating / opening a database file

```{r}
con <- dbConnect(duckdb(), dbdir = "my_data.duckdb", read_only = FALSE)
```

-   Writing the municipality data to DuckDB as tables.

```{r}
dbExecute(con, "INSTALL spatial; LOAD spatial;")

# List of municipality object names already in R
muni_names <- c(
  "parcels_break_o_day",
  "parcels_brighton",
  "parcels_burnie",
  "parcels_central_coast",
  "parcels_central_highlands",
  "parcels_circular_head",
  "parcels_clarence",
  "parcels_derwent_valley",
  "parcels_devonport",
  "parcels_dorset",
  "parcels_flinders",
  "parcels_george_town",
  "parcels_glamorgan_spring_bay",
  "parcels_glenorchy",
  "parcels_hobart",
  "parcels_huon_valley",
  "parcels_kentish",
  "parcels_kingborough",
  "parcels_king_island",
  "parcels_latrobe",
  "parcels_launceston",
  "parcels_meander_valley",
  "parcels_northern_midlands",
  "parcels_sorell",
  "parcels_southern_midlands",
  "parcels_tasman",
  "parcels_waratah_wynyard",
  "parcels_west_coast",
  "parcels_west_tamar"
)

# Loop through each object, convert + write to DuckDB
for (muni in muni_names) {
  message("Processing: ", muni)
  
  # Get the sf object from environment
  sf_obj <- get(muni, envir = .GlobalEnv)
  
  # Convert to plain data frame with WKB geometry
  df <- sf_obj |>
    st_drop_geometry() |>
    as.data.frame()
  df$geometry <- st_as_binary(sf_obj$geometry)
  
  # Write to DuckDB with same name as in R
  dbWriteTable(con, muni, df, overwrite = TRUE)
}

```

-   Writing the Address Master and property listing data to DuckDB as tables.

```{r}
# Build SQL string with UNION ALL
sql_parts <- lapply(muni_names, function(tbl) {
  muni_name <- toupper(gsub("^parcels_", "", tbl))   # e.g. "parcels_hobart" -> "HOBART"
  
  paste0(
    "SELECT PID, * EXCLUDE(PID), '", muni_name, "' AS MUNICIPALITY ",
    "FROM ", tbl, " ",
    "WHERE PID <> 0"
  )
})

master_sql <- paste(sql_parts, collapse = "\nUNION ALL\n")

# Wrap into CREATE TABLE
create_master_sql <- paste0(
  "CREATE OR REPLACE TABLE parcels_master AS\n",
  master_sql,
  ";"
)

# Run in DuckDB
dbExecute(con, create_master_sql)
```

-   Checking the data in DuckDB

```{sql, connection=con}
select count(*) from parcels_master;
select count(distinct(PID)) from parcels_master where TENURE_TY !='Crown Land';
```

```{sql, connection=con}
select * from parcels_master limit 5;
```

-   Addding index and set geometry column as binary.

```{r}
# `{sql, connection=con}Explicitly coerce to data.frame after dropping geometry
address_master_tbl <- Address_Master |>
  st_drop_geometry() |>
  as.data.frame()

# Then add a WKB geometry column from the original sf object
address_master_tbl$geometry <- st_as_binary(Address_Master$geometry)


dbWriteTable(con, "address_master",address_master_tbl, overwrite = TRUE)

dbWriteTable(con, "property_listing_domain", property_listing_domain, overwrite = TRUE)

# (optional) useful indexes
dbExecute(con, "CREATE INDEX IF NOT EXISTS idx_parcels_pid ON parcels_master(PID)")
dbExecute(con, "CREATE INDEX IF NOT EXISTS idx_parcels_muncipality ON parcels_master(MUNICIPALITY)")
dbExecute(con, "CREATE INDEX IF NOT EXISTS idx_addr_pid    ON address_master(PID)")
dbExecute(con, "CREATE INDEX IF NOT EXISTS idx_listing_pid ON property_listing_domain(full_address)")
```

-   Checking for duplicate PIDs in parcels_master

```{sql, connection=con}

select distinct(PID), count(*) from parcels_master group by PID having count(*)>1 limit 5;


select * from parcels_master where PID='6585214';
```

-   Dropping the view if it exists.

```{sql, connection=con}
drop view if exists PARCEL_ADDRESS_MASTER;
```

-   creating a view that joins parcels_master and address_master on PID to create a comprehensive parcel address master view.
-   we eliminate P.PID=0 because it is either a water body or any kind of road.

```{sql, connection=con}
CREATE VIEW PARCEL_ADDRESS_MASTER AS select distinct(P.PID
),P.MUNICIPALITY, concat(
    cast(A.ST_NO_FROM as varchar), ' ', A.STREET, ' ', A.ST_TYPE, ' ', A.LOCALITY, ' TAS ', A.POSTCODE, ' Australia'
) as FULL_ADDRESS,ANY_VALUE(P.geometry) as parcel_geometry,P.POT_PID, ANY_VALUE(A.geometry) as address_geometry
from parcels_master P inner join address_master A on P.PID = A.PID where P.PID !='0' group by P.PID,P.MUNICIPALITY, FULL_ADDRESS,P.POT_PID;
```

```{sql, connection=con}
select * from PARCEL_ADDRESS_MASTER where PID='6003587';
```

-   Joining the property listing data with the parcel address master view on full_address to get a combined dataset with parcel and address geometries.

```{sql, connection=con}
SELECT p.PID ,p.POT_PID,d.full_address,d.latitude,d.longitude,ST_AsText(ST_GeomFromWKB(p.parcel_geometry))   AS parcel_wkt,
  ST_AsText(ST_GeomFromWKB(p.address_geometry))  AS address_wkt
FROM property_listing_domain d
JOIN PARCEL_ADDRESS_MASTER p
  ON upper(trim(replace(d.full_address, chr(160), ' '))) =
     upper(trim(replace(p.full_address, chr(160), ' ')));
```

```{sql, connection=con}
select * from PARCEL_ADDRESS_MASTER where PID='5273454';
```

```{r}
# `con` is your DBI connection (already open)
dbWriteTable(con, "tesm_majority", majority_tab_with_pid, overwrite = TRUE)
```

```{sql, connection=con}
select * from tesm_majority limit 5;
```

```{sql, connection=con}
select p.PID,FULL_ADDRESS,CLASS_LABEL,PIXELS,ST_AsText(ST_GeomFromWKB(p.parcel_geometry))   AS parcel_wkt from tesm_majority m join PARCEL_ADDRESS_MASTER p on m.PID=p.PID;
```
