---
title: "Literature Review Instructions"
author:
- name: Rafli Ramdani
  degrees: TBA
  email: rram0049@student.monash.edu
- name: Azham Hassan Venghat
  degrees: TBA
  email: aven0024@student.monash.edu
phone: TBA
email: BusEco-Econometrics@monash.edu
organization: TBA
bibliography: references.bib
format: report-pdf
---


# Objective (< 300 words)

To support the development of a flexible, high‑resolution modelling framework by conducting a structured literature review of:

1. The **existing Tasmanian Enterprise Suitability Maps (TESMs)** – their inputs, methods, and limitations  (extensive, around 4500 words)

Definition:

Inputs: 

Methods:

Limitations:

2. Comparable **crop suitability models** from other regions



3. **Emerging, high‑value, or climate‑resilient crops** suitable for Tasmania’s cool temperate climate  

Emerging: Walnuts

High-value: Apples

Climate-resilient: Oats/triticale, Canola


Your review will guide improvements to TESMs and the inclusion of specialty crops (e.g. truffles).

---

# Part 1: Review of the Existing Tasmanian Enterprise Suitability Maps (TESMs)

## Tasks

- **Summarise current model features**  
  - List/tabulate all crops currently included  
  Ans: combining steps documented on `all_crops_combining.R` file.
  
```{r, warning=F, message=F}
library(tidyverse)

crops_all <- readRDS("data/all_crops_df.rds")

head(crops_all,5)
```

```{r}
distinct(crops_all$`Crop Type`)
```


  - Identify spatial inputs (e.g. rainfall, soil type, temperature, slope)

```{r}
glimpse(crops_all)
str(crops_all, width=30)
colnames(crops_all)
```

  - Describe modelling approach (deterministic rules, scoring thresholds)  


  - Note assumptions (irrigation assumed, no uncertainty, uniform management)  
    
    - Irrigation assumed: ....
    
    - No uncertainty: ...
    
    - Uniform Management: ...


- **Critically assess limitations**  
  - Missing inputs (e.g. chill hours, pH, drainage, frost exposure)
  
  - Lack of uncertainty estimation or validation  
  
  - Inflexibility for adding niche or specialty crops  
  
  - Over generalization (ignoring micro climates, management practices) 
  

## Suggested Comparator Models

Briefly describe **2–3** of the following alternative suitability frameworks, focusing on inputs, methods, uncertainty, extensibility:

- **New Zealand S‑map & Crop Suitability Layers** (Manaaki Whenua – Landcare Research)  
  - Includes truffles and other specialty crops  
  - Integrates soil, climate & management  
  - Modular and extensible design

- Victorian Horticulture Crop Suitability Framework  

- Queensland Land Suitability Guidelines (DES)  

- FAO EcoCrop / GAEZ  

- USDA Crop Suitability Tools  

- CSIRO land evaluation frameworks  



---

# Part 2: Literature Review on Emerging & Climate‑Resilient Crops

## Selection Criteria

Identify crops that are:

- **Not** in TESMs  
- High‑value or specialty (e.g. truffles, saffron, hops) 

  - Ginseng
  
  - Apples
  
  - 

- Climate‑resilient (drought/frost/salinity tolerance)  

  - Oats/Triticale (frost, less water)
  
  - ...  (high salinity)
  
  - Sorghum (drought)
  
  - Quinoa (drought)


- Suitable for regenerative, small‑scale systems in Tasmania  

  - Walnuts
  
  - 
  
  - 
  

- Under trial or limited local production (bonus)


## Crop Summary Requirements

For each candidate crop, capture at minimum the details required as input to the existing TESM, and any others indentified from your literature review, e.g:

- **Climatic**: temperature range, rainfall, chill hours, frost/heat tolerance  
- **Soil**: texture, drainage, pH, organic matter  
- **Topography**: slope, elevation suitability  
- **Water**: irrigation needs or dryland tolerance  
- **Management**: pest/disease issues, growing season, pollination  
- **Markets**: value, demand trends, Tasmanian viability  
- **Model Inputs**: spatial variables you could add to TESMs  

---

# Deliverables

1. **Narrative Report** (4–6 pages)  
   - Executive summary  
   - TESMs overview & limitations  
   - Comparator model insights  
   - 5–10 recommended crops (with focus on truffles)  
   - Recommendations for new inputs & uncertainty handling  

2. **Reference List**  
Please only use Peer‑reviewed journals, government & industry reports. Use reference formating according to Monash unit guidelines.  


# Resources

- **TESMs**: <https://dpipwe.tas.gov.au/agriculture/land-resources/land-capability-and-suitability/enterprise-suitability>  
- **NZ S‑map**: <https://smap.landcareresearch.co.nz/>  
- **FAO EcoCrop**: <https://ecocrop.fao.org/>  
- CSIRO, AgriFutures, DPIPWE, AgVic, QLD DES publications  
- Google Scholar, Scopus, Web of Science, Trove  

---

